Ice sheet mass flows

### 2024-07-28 ms.tex
Writing

  Abstract 1:

  We combine recent estimates of ice mass change that include ice at or near flotation (i.e., terminus change in Greenland) with grounded ice mass balance to produce the first total mass balance product for the Greenlandic and Antarctic ice sheets. Greenland mass loss is 30 \% higher than previous estimates due to terminus retreat. This additional mass loss does not contribute significantly to sea level rise as it is already at or near flotation.

  Abstract 2:

  Understanding the processes involved in ice sheet mass loss requires knowledge production and sharing, in addition to field observations and computer modeling. Tabular data is optimized for computers and useful for humans, but graphical presentations can provide significantly higher information density. Here we present several Sankey diagrams depicting all mass flow of water (in various phases) in Greenland and Antarctica. These diagrams show mass loss in Greenland is 330 Gt yr$^{-1}$  which is 30 \% higher previous estimates that neglect terminus retreat of ice at or near flotation.

\end{abstract}
\end{frontmatter}

@@

Mass flows of ice sheets, and subsequent mass gains and losses, is an of-studied topic in the domain of glaciology. However, few studies consider all processes and their relative magnitudes and uncertainties - this is typically limited to review papers, which may not focus on quantitative assessment of each process.

@@

\begin{figure*}
\centering{\includegraphics[width=0.85\textwidth]{fig.png}}
\caption{Sankey mass flow diagrams for Antarctica and Greenland, and Antarctica split into East, West, and Peninsula. All widths are proportional within and between images. Because Sankey diagrams balance all inputs and outputs, mass losses require a `drawdown' input (red) to balance the larger outputs, and mass gains requires an `accumulation' output (black) to balance the larger inputs.}
\label{fig}
\end{figure*}

@@

  The flows can be somewhat tied to physical space as in the Minard map mentioned above, but are not done so explicitly here. An similar display to the diagrams presented here, of glacier mass flow by \citet[Fig. 2]{cogley_2011}, shows flow overlaid an a glacier schematic, but with neither flow magnitude nor proportionality. Nonetheless, there is an implicit physical connection here for most flow. Inputs all come from the atmosphere (with the exception of drawdown and frontal advance), and outputs pointing down are all mass flow into the ocean.

@@

We use existing published values, some new estimates, and some derived values (See Tables \ref{tab:gl} and \ref{tab:aq} in Appendix A).

@@

Here we describe our new estimates.

@@

@@

>>>>

  Understanding the processes involved in ice sheet mass loss requires knowledge production and sharing, in addition to field observations and computer modeling. Tabular data is optimized for computers and useful for humans, but graphical presentations can provide significantly higher information density. Here we present several Sankey diagrams depicting all mass flow of water in various phases in Greenland and Antarctica. Sankey diagrams help bring perspective to the interplay and relative importance of snowfall, rain, drawdown, refreezing, submarine melt, iceberg calving, and grounding line retreat in contributing to ice sheet mass balance.

  This work quantifies both grounded and floating ice mass loss. From this, mass loss in Greenland is 330 Gt yr$^{-1}$ which is \(\sim\)25 \% higher than the \(\sim\)270 Gt yr$^{-1}$ estimates that neglect terminus retreat of ice at or near flotation. Mass loss in Antarctica is 275 Gt yr$^{-1}$ which is \(\sim\)85 \% higher than the \(\sim\)150 Gt yr$^{-1}$ estimates of grounded ice mass loss.

  This work also reports gross not net values, which is useful for example for quantifying total freshwater export from an ice sheet. For example, freshwater export due to frontal retreat is not 60 Gt yr$^{-1}$, but rather nearly six times larger at 345 Gt yr$^{-1}$ due to 345 Gt yr$^{-1}$ of frontal retreat offset by 285 Gt yr$^{-1}$ of frontal advance, with the difference as the commonly reported value of 60 Gt yr$^{-1}$.
\end{abstract}
\end{frontmatter}

@@

Mass loss from the grounded portion of the Greenlandic and Antarctic ice sheets contributes to sea level rise (SLR), and can be captured by a single value per ice sheet. Reporting a single value is beneficial because it simplifies interpretation and comparison. However, ice sheets are complex systems and while their contribution to sea level rise is important, it is not a holistic view and does not capture the general health and state of the ice sheet. Few studies consider all processes and their relative magnitudes and uncertainties - this is typically limited to review papers, which may not focus on quantitative assessment of each process. As one example, an ice sheet that increases output via discharge or submarine melting by X \% but has that offset by an equal increase in snowfall would report no net mass change or net SLR contribution, but has entered a different state when viewing constituent terms or gross values.

@@

\begin{figure*}
\centering{\includegraphics[width=0.85\textwidth]{fig_aq_gl.png}}
\caption{Sankey mass flow diagrams for Antarctica and Greenland, and Antarctica split into East, West, and Peninsula. All widths are proportional within and between images. Because Sankey diagrams balance all inputs and outputs, mass losses require a `drawdown' input (red) to balance the larger outputs, and mass gains requires an `accumulation' output (black) to balance the larger inputs.}
\label{fig}
\end{figure*}

@@

A similar display to the diagrams presented here by \citet[Fig. 2]{cogley_2011} shows flows overlaid an a glacier schematic and here we build on that work by adding magnitude of processes and making the graphics proportional to magnitudes.

@@

We use a combination of existing published values, new estimates, and derived values.

Existing published values include

@@

@@

\subsection{Sankey diagrams}

The Sankey diagrams shown here are generated from a script that combines a CSV file of values with a \LaTeX\ template that uses the TikZ Sankey package \citep{sankey}. This architecture makes it trivial to generate similar diagrams for other time periods, differences between time periods, after generating CSV files with representative values.

@@

\section{Appendix B}
\label{sec:appendix:B}

\begin{figure*}
\centering{\includegraphics[width=0.85\textwidth]{fig_aq_parts.png}}
\caption{Sankey mass flow diagrams for Antarctica and Greenland, and Antarctica split into East, West, and Peninsula. All widths are proportional within and between images. Because Sankey diagrams balance all inputs and outputs, mass losses require a `drawdown' input (red) to balance the larger outputs, and mass gains requires an `accumulation' output (black) to balance the larger inputs.}
\label{fig}
\end{figure*}
